---
title: "PuMBA Demonstration"
always_allow_html: true
output:
  github_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
    toc_float:
      toc_collapsed: yes
---

# Aims

Here I will show the basic usage of PuMBA on a couple of simulated datasets.  I will first go through the data briefly, and then demonstrate fitting the model to this data.


# Libraries

```{r, message=FALSE}
library(tidyverse)
library(brms)
library(skimr)
library(ggbeeswarm)
library(job)
library(broom)
library(broom.mixed)
library(lme4)
library(lmerTest)

theme_set(theme_light())
set.seed(555)
```


# TAC Data: 2TC

I will first demonstrate the use of PuMBA using a simulated dataset taken from our first paper on these types of models, called SiMBA.  The dataset comes from [here](https://github.com/mathesong/SiMBA_Materials).


## Exploring the data

### Loading

I saved the simulated sample data in the RawData folder.

```{r}
simdat <- readRDS("../RawData/simdata.rds")

head(simdat)

skim(simdat)
```


### Measured Data

This dataset consists of 10 controls and 10 patients.  The sigma value is set to about 10% of the mean TAC value. There is a group difference of 20% (Cohen's d = 0.55) in all regions.  The unique identifier for each individual is called `WAYID`.


#### TAC

For now, I'll just work with the first subject and one region.

```{r}
subjdat <- simdat %>% 
  filter(WAYID==WAYID[sample(1:nrow(simdat), 1)] & Region=="dor")
```

Let's have a look

```{r}
subjdat %>% 
  select(t_tac, Measured = TAC, True = TAC_true) %>% 
  ggplot(aes(x=t_tac, y=Measured)) +
  geom_point() + 
  geom_line(aes(y=True), linetype="dashed") +
  labs(title="Simulated TAC from the DLPFC",
       subtitle="Points represent the measured TAC, and the dashed line represents the true TAC without error",
       x = "Time (min)",
       y = "Radioactivity")
```





#### Arterial Input function

The AIF is modelled with a tri-exponential function.

```{r}
predict_aif <- function(time, t0, slope, peaktime, A, B, C, alpha, beta, gamma) {
  
  t_corr <- time - t0
  
  if(t_corr <= 0) {
    pred <- 0
    return(pred)
  }
  
  if(t_corr <= peaktime) {
    pred <- t_corr * slope
    return(pred)
  }
  
  if(t_corr > peaktime) {
    pred <- A*exp(-alpha*t_corr) + 
      B*exp(-beta*t_corr) + 
      C*exp(-gamma*t_corr)
    return(pred)
  }
}

time <- seq(0, 105, by=0.01)
aif <- map_dbl(time, ~with(subjdat[1,], 
            predict_aif(.x, t0, slope, peaktime, A, B, C, alpha, beta, gamma)))

AIFdat <- tibble(
  time = time, 
  AIF = aif
)

ggplot(AIFdat, aes(x=time, y=AIF)) +
  geom_line(colour="red")
```



So now, we can show the AIF and the TAC together.

```{r}
subjdat %>% 
  select(t_tac, Measured = TAC, True = TAC_true) %>% 
  ggplot(aes(x=t_tac, y=Measured)) +
  geom_point() + 
  geom_line(aes(y=True), linetype="dashed") +
  labs(title="Simulated TAC and AIF",
       subtitle="Black represents the TAC and true TAC, while red represents the AIF",
       x = "Time (min)",
       y = "Radioactivity") +
  geom_line(data = AIFdat, aes(x=time, y=AIF), colour="red") +
  coord_cartesian(ylim = c(-0.02, 0.3))
```

#### Whole Blood

We did not have access to whole blood data, as this was not measured. However, we did measure whole plasma data, which we substitute for whole blood data. In preprocessing, I fit a spline to this data, and calculated the mean blood concentration during each frame. This allows us to input blood concentration as a vector like the TAC.

```{r}
ggplot(subjdat, aes(x=t_tac, y=TAC)) +
  geom_point() + 
  geom_line(aes(x=t_tac, y=meanblood), colour="blue") +
  geom_point(aes(x=t_tac, y=meanblood), colour="blue") +
  geom_line(data = AIFdat, aes(x=time, y=AIF), colour="red") +
  labs(title="Blood, AIF and TAC",
       subtitle="The whole blood is represented in blue, the AIF in red, and the TAC in black",
       x = "Time (min)",
       y = "Radioactivity")
```



### Simulation Data

Now, let's explore some of the characteristics of the simulation. 

#### Measurement Error

First, let's look at measurement error.  This is simulated from the fitted spline.  Let's examine the values of the measurement error SD over the course of the single TAC above.

```{r}
subjdat %>% 
  mutate(Sigma_Percentage = 100*(exp(sigwig))) %>% 
  ggplot(aes(x=t_tac, y=Sigma_Percentage)) +
  geom_point() +
  geom_line() +
  geom_hline(aes(yintercept=100), linetype="dashed") +
  labs(x="Time (min)",
       y="Measurement Error (%)",
       title="Measurement error function",
       subtitle="Measurement error as a percentage of the mean value across the PET measurement")
```

This function is partially related to the durations of the frames, but not completely.

```{r}
ggplot(data=subjdat, aes(x=t_tac, y=durations)) +
  geom_point() +
  geom_line() +
  labs(x="Time (min)",
       y="Frame Duration (min)",
       title="Frame durations",
       subtitle="Frame durations over the course of the PET measurement")
```


The true sigma at each point in time is taken from the addition of the true sigma for each ROI and individual, `sigma_true`, and adding `sigwig` (the sigma wiggly function), to make `sigma_w_true`.

Let's show this represented around the TAC.

```{r}
subjdat %>% 
  select(t_tac, Measured = TAC, True = TAC_true, sigma = sigma_w_true) %>% 
  mutate(sigma = exp(sigma)) %>% 
  ggplot(aes(x=t_tac, y=Measured)) +
  geom_point(size=3, aes(y=True)) + 
  geom_errorbar(aes(ymax = True + 1.96*sigma, ymin = True - 1.96*sigma)) +
  geom_point(colour="red", size=3, shape=8) + 
  geom_line(aes(y=True), linetype="dashed") +
  labs(title="True and measured TACs and measurement error",
       subtitle="True TAC and the measurement error in black, with the measured TAC in red",
       x = "Time (min)",
       y = "Radioactivity")
```

#### Parameters by Region

The true parameters are included as simulated from the model.  Here I'll make a data frame with only the parameters.

```{r}
pardat <- simdat %>% 
  filter(!duplicated(paste0(WAYID, Region))) %>% 
  select(WAYID, Region, contains("_true"), Group) %>% 
  select(-TAC_true, -sigma_w_true) %>% 
  gather(Parameter, Value, -WAYID, -Region, -Group) %>% 
  mutate(Value = exp(Value),
         Parameter = str_remove(Parameter, "log"),
         Parameter = str_remove(Parameter, "_true"),
         Parameter = fct_inorder(Parameter))

pardat_wide <- simdat %>% 
  filter(!duplicated(paste0(WAYID, Region))) %>% 
  select(WAYID, Region, contains("_true"), Group) %>% 
  select(-TAC_true, -sigma_w_true)
```

Let's look at the differences between true parameters in the different regions.

```{r, fig.width=10, fig.height=10}
ggplot(pardat, aes(x=Value, colour=Region, fill=Region)) + 
  geom_histogram(alpha=0.4) +
  facet_grid(Region ~ Parameter, scales="free") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
```


Note the regions are as follows:

* acn : anterior cingulate cortex
* amy : amygdala
* cin : posterior cingulate cortex
* dor : dorsolateral prefrontal cortex
* hip : hippocampus
* ins : insula
* med : medial prefrontal cortex
* pip : parahippocampal cortex
* WAY_rphelpscs_c : dorsal raphe nucleus


#### Parameters by Group

There are only group differences in BP~ND~, which are the same across all regions: 20%. Let's take a look by looking at only the DLPFC ROI.


```{r}
parroidat <- pardat %>% 
  filter(Region=="dor")
```

Let's look at the differences between true parameters in the different regions.

```{r}
ggplot(parroidat, aes(x=Group, y=Value, colour=Group, fill=Group)) + 
  geom_violin(alpha=0.4) +
  geom_beeswarm() +
  facet_wrap(~Parameter, scales="free")
```

This really speaks to the difficulty of sampling variation... Really hard to see anything at all there in BP~ND~ in this particular sample.


## NLS Modelling

We would usually model these TACs using nonlinear least squares, and then perform statistical inference.  Let's do that for comparison here.  First, we need to define the model in R code.

```{r}
two_compartment_log = function(logk1, logvnd, logbpnd, logk4, logvb, time, 
                               t0_aif, b_aif, 
                               lambda1_aif, lambda2_aif, lambda3_aif, 
                               A1_aif, A2_aif, A3_aif, tstar_aif, 
                               bloodval){
  
  
  k1 = exp(logk1)
  vnd = exp(logvnd)
  bpnd = exp(logbpnd)
  k4 = exp(logk4)
  vb = exp(logvb)
  
  k2 = k1 / vnd
  k3 = bpnd * k4
  
  R1 = 0.5 * (k2 + k3 + k4 + sqrt((k2 + k3 + k4)^2 - 4 * k2 * k4))
  R2 = 0.5 * (k2 + k3 + k4 - sqrt((k2 + k3 + k4)^2 - 4 * k2 * k4))
  L1 = (k1 * (R1 - k3 - k4))/(R1 - R2)
  L2 = (k1 * (k3 + k4 - R2))/(R1 - R2)
  
  time <- time - t0_aif
  tstar_aif <- tstar_aif - t0_aif
  
  out <- (time>0)*((1-vb)*
                     
   ((time<tstar_aif)*(b_aif*L1*(exp(-R1*time)/R1^2 + time/R1 - 1/R1^2) +
                        b_aif*L2*(exp(-R2*time)/R2^2 + time/R2 - 1/R2^2)) +
      
      (time>=tstar_aif)*(b_aif*L1*exp(-R1*time)*(tstar_aif/R1*exp(R1*tstar_aif) - 
                                                   1/R1^2*exp(R1*tstar_aif) + 
                                                   1/R1^2) + 
         L1*A1_aif*(exp(-lambda1_aif*time)/(R1-lambda1_aif) -
                      exp(R1*tstar_aif-R1*time-lambda1_aif*tstar_aif)/(R1-lambda1_aif)) +
         L1*A2_aif*(exp(-lambda2_aif*time)/(R1-lambda2_aif) -
                      exp(R1*tstar_aif-R1*time-lambda2_aif*tstar_aif)/(R1-lambda2_aif)) +
         L1*A3_aif*(exp(-lambda3_aif*time)/(R1-lambda3_aif) -
                      exp(R1*tstar_aif-R1*time-lambda3_aif*tstar_aif)/(R1-lambda3_aif)) +
         b_aif*L2*exp(-R2*time)*(tstar_aif/R2*exp(R2*tstar_aif) - 
                                   1/R2^2*exp(R2*tstar_aif) + 1/R2^2) +
         L2*A1_aif*(exp(-lambda1_aif*time)/(R2-lambda1_aif) -
                      exp(R2*tstar_aif-R2*time-lambda1_aif*tstar_aif)/(R2-lambda1_aif)) +
         L2*A2_aif*(exp(-lambda2_aif*time)/(R2-lambda2_aif) -
                      exp(R2*tstar_aif-R2*time-lambda2_aif*tstar_aif)/(R2-lambda2_aif)) +
         L2*A3_aif*(exp(-lambda3_aif*time)/(R2-lambda3_aif) -
                      exp(R2*tstar_aif-R2*time-lambda3_aif*tstar_aif)/(R2-lambda3_aif)))) + 
                     vb*bloodval)
  
  out <- as.numeric(out)
  
  return(out)
}
```

Now, we calculate model weights.

```{r}
nls_simdat_wide <- simdat %>% 
    pivot_wider(names_from = Region, values_from = TAC, id_cols = c(WAYID, t_tac, durations)) %>% 
    group_by(WAYID) %>% 
    nest() %>% 
    mutate(weights = map(data, ~kinfitr::weights_create(
      t_start = .x$t_tac - 0.5*.x$durations,
      t_end   = .x$t_tac + 0.5*.x$durations,
      tac = abs(.x$dor), radioisotope = "C11"
    ))) %>% 
    unnest(c(data, weights)) %>% 
    pivot_longer(cols = dor:WAY_rphelpscs_c, names_to = "Region", values_to = "TAC") %>% 
    ungroup()
```

... and then bind them to the data

```{r}
nls_simdat <- left_join(nls_simdat_wide, simdat, by=c("WAYID", "t_tac", 
                                                    "durations", "Region", 
                                                    "TAC")) %>% 
    group_by(WAYID, Region) %>% 
    nest() %>% 
    mutate(weightsvals = map(data, ~pull(.x, weights)))
```

Now we're ready to fit the TACs.  I fit them with multiple starting values (10 fits per TAC), to try to avoid falling into a local minimum.

```{r}
nls_simdat <- nls_simdat %>%
  mutate(fit = map(data, ~nls.multstart::nls_multstart(
    TAC ~ two_compartment_log(logk1, logvnd, logbpnd, logk4, logvb,
                              t_tac, t0, slope,
                              alpha, beta, gamma,
                              A, B, C, peaktime,
                              meanblood), data = .x,
    start_lower = c(logk1=-3, logvnd=-1.5, logbpnd=1.5, logk4=-4.5, logvb=-4.5),
    start_upper = c(logk1=-2, logvnd=-0.5, logbpnd=2.5, logk4=-3.5, logvb=-3.5),
    upper = c(logk1=0, logvnd=1, logbpnd=3.5, logk4=-1, logvb=-2.5),
    lower = c(logk1=-6, logvnd=-4, logbpnd=-1, logk4=-4, logvb=-7),
    modelweights = weights, iter=10, supp_errors = "Y")))
```

Now, let's extract the values

```{r}
nls_simres <- nls_simdat %>% 
    mutate(coef = map(fit, ~as.data.frame(as.list(coef(.x))))) %>% 
    ungroup() %>% 
    unnest(coef) %>% 
    mutate(logvt = log( exp(logvnd)*(1 + exp(logbpnd)) ),
           logbpp = log( exp(logvnd)*(exp(logbpnd)) ),
           Group = map_chr(data, ~.x$Group[1])) %>% 
    select(WAYID, Region, logk1:Group)
  
truevals <- nls_simdat %>% 
    select(WAYID, Region, data) %>% 
    unnest(data) %>% 
    filter(!duplicated(paste0(Region, WAYID))) %>% 
    select(WAYID, Region, ends_with("_true")) %>% 
    select(-TAC_true) %>% 
    mutate(logVt_true = log( exp(logVnd_true)*(1 + exp(logBPnd_true)) ),
           logBPp_true = log( exp(logVnd_true)*(exp(logBPnd_true))) )
  
nls_simres <- nls_simres %>% 
    left_join(truevals)
```


### Comparison to true values

Let's see how well we've estimated the binding. 

First, for BP~P~:

```{r}
ggplot(nls_simres, aes(x=logBPp_true, y=logbpp, colour=Region)) + 
  geom_point() +
  facet_wrap(~Region, scales="free") + 
  geom_smooth(method="lm", se=FALSE)

nls_simres %>% 
  group_by(Region) %>% 
  summarise(cor = cor(logBPp_true, logbpp))
```


And then for BP~ND~:

```{r}
ggplot(nls_simres, aes(x=logBPnd_true, y=logbpnd, colour=Region)) + 
  geom_point() +
  facet_wrap(~Region, scales="free") + 
  geom_smooth(method="lm", se=FALSE)

nls_simres %>% 
  group_by(Region) %>% 
  summarise(cor = cor(logBPnd_true, logbpnd))
```


It was for this reason that we opted to use BP~P~ for PuMBA rather than BP~ND~.

## Conventional Statistical Analysis

Next, we move to statistical comparison.

### t-tests

For the comparisons below, the difference is calculated as Group1 - Group2, so it is negative. I will correct this so that they are positive.


```{r}
nls_simres_t_bpp <- nls_simres %>% 
  group_by(Region) %>% 
  nest() %>% 
  mutate(res = map(data, ~broom::glance(t.test(logbpp ~ Group, data=.x)))) %>% 
  select(Region, res) %>% 
  unnest(res) %>% 
  mutate(across(.cols=c(estimate, conf.low, conf.high), .fns = ~.x*-1))

nls_simres_t_bpp

ggplot(nls_simres_t_bpp, aes(x=Region, y=estimate)) +
  geom_point() + 
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high)) +
  geom_hline(yintercept=0.182, colour="red", linetype="dashed") +
  coord_flip()
```

These are mostly in the right direction, but the 95% CIs around the estimates are very wide indeed...


### LME

In order to improve our estimation, we can use linear mixed effects models.

```{r}
nls_simres_lme_bpp <- broom.mixed::tidy(lmerTest::lmer(logbpp ~ 1 + Region + Group + (1|WAYID), 
                                                  data=nls_simres))

nls_simres_lme_bpp

nls_simres_lme_bpp %>% 
  filter(term=="GroupPatient") %>% 
  ggplot(aes(x=term, y=estimate)) +
  geom_point() + 
  geom_errorbar(aes(ymin = estimate + 1.96*std.error, 
                    ymax = estimate - 1.96*std.error)) +
  geom_hline(yintercept=0.182, colour="red", linetype="dashed") +
  coord_flip()
```


This results in tigher confidence intervals, which is a good sign!


## PuMBA

Now, we would like to run PuMBA on this data. 


This first requires making sure that the dummy variables work correctly. We'll make Region into a factor.

```{r}
pumba_modeldat <- nls_simres %>% 
  mutate(Region = fct_inorder(Region)) %>% 
  rename(
    logK1 = logk1,
    logVnd = logvnd,
    logBPnd = logbpnd,
    logBPp = logbpp,
    logk4 = logk4) %>% 
  rename(ID = WAYID)
```


Then we define our prior

```{r}
pumba_prior <- c(
  set_prior("normal(-2.5, 0.25)", resp = "logK1"),
  set_prior("normal(-1, 0.25)", resp = "logVnd"),
  set_prior("normal( 1, 0.25)", resp = "logBPp"),
  set_prior("normal(-4, 0.25)", resp = "logk4"),
  
  set_prior("normal(0, 0.3)", resp = "logK1", class = "sd", group="ID"),
  set_prior("normal(0, 0.3)", resp = "logVnd", class = "sd", group="ID"),
  set_prior("normal(0, 0.3)", resp = "logBPp", class = "sd", group="ID"),
  set_prior("normal(0, 0.3)", resp = "logk4", class = "sd", group="ID"),
  
  set_prior("normal(0, 0.1)", resp = "logK1", class = "sigma"),
  set_prior("normal(0, 0.1)", resp = "logVnd", class = "sigma"),
  set_prior("normal(0, 0.1)", resp = "logBPp", class = "sigma"),
  set_prior("normal(0, 0.1)", resp = "logk4", class = "sigma"),
  
  set_prior("normal(0, 0.1)", resp = "logVnd", class = "sd", group="Region"),
  set_prior("normal(0, 0.1)", resp = "logk4", class = "sd", group="Region"),
  
  set_prior("normal(0, 0.3)", coef="Regionacn", resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regionamy", resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regioncin", resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regionhip", resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regionins", resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regionmed", resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regionpip", resp="logK1"),
  set_prior("normal(0, 0.3)", coef="RegionWAY_rphelpscs_c", resp="logK1"),
  
  set_prior("normal(0, 0.3)", coef="Regionacn", resp="logBPp"),
  set_prior("normal(0, 0.3)", coef="Regionamy", resp="logBPp"),
  set_prior("normal(0, 0.3)", coef="Regioncin", resp="logBPp"),
  set_prior("normal(0, 0.3)", coef="Regionhip", resp="logBPp"),
  set_prior("normal(0, 0.3)", coef="Regionins", resp="logBPp"),
  set_prior("normal(0, 0.3)", coef="Regionmed", resp="logBPp"),
  set_prior("normal(0, 0.3)", coef="Regionpip", resp="logBPp"),
  set_prior("normal(0, 0.3)", coef="RegionWAY_rphelpscs_c", 
            resp="logBPp"),
  
  set_prior("normal(0, 0.2)", coef="GroupPatient", 
            resp="logBPp"),

  set_prior("lkj(1)", class="cor", group = "ID"),
  set_prior("lkj(2)", class="cor", group = "Region"),
  set_prior("lkj(2)", class="rescor"))

```

... and our formula

```{r}
bf_K1 <- bf(logK1 ~ Region + (1|k|ID), center = F)
bf_Vnd <- bf(logVnd ~ 1 + (1|j|Region) + (1|k|ID), center = F)
bf_BPp <- bf(logBPp ~ 1 + Region + Group + (1|k|ID), 
              center = F)
bf_k4 <- bf(logk4 ~ 1 + (1|j|Region) + (1|k|ID), center = F)

pumba_formula <- brms::mvbrmsformula(bf_K1 + bf_Vnd + bf_BPp + bf_k4 + set_rescor(TRUE))
```

... and we can fit the model:

```{r}
pumba_fit <- brm(pumba_formula,
    family=gaussian(),
    data = pumba_modeldat,
    prior = pumba_prior,
    backend = "cmdstanr", 
    inits = 0, cores=4, 
    iter=4000,
    silent = 2, refresh=0)

pumba_fit
```




## Comparison of estimates

Now, let's just compare the model estimates and their 95% intervals from all the different tests.  First, I'll just extract the estimates and put them all together.  I'll also add the SiMBA results for comparison.

```{r, fig.width=8}
simcompare_t <- nls_simres_t_bpp %>% 
  select(Region, estimate, conf.low, conf.high) %>% 
  mutate(Parameter = "BPND") %>% 
  mutate(Parameter = "BPP") %>% 
  mutate(Test = paste0("t-test")) %>% 
  rename(U95 = conf.low, L95 = conf.high) %>% # reversing because we
                                              # multiplied by -1 earlier
  mutate(SE = (U95 - estimate)/1.96 )

simcompare_lme <- nls_simres_lme_bpp %>% 
  filter(term == "GroupPatient") %>% 
  mutate(Parameter = "BPP",
         Region = "All Regions",
         Test = paste0("LME"),
         conf.low = estimate + 1.96*std.error,
         conf.high = estimate - 1.96*std.error) %>%
  select(Region, estimate, conf.low, conf.high, SE = std.error, 
         Parameter, Test) %>% 
  rename(L95 = conf.low, U95 = conf.high)
  
simcompare_pumba <- as_tibble(rstan::summary(pumba_fit$fit)$summary, 
                           rownames = "Parameter") %>% 
  filter(str_detect(Parameter, "GroupPatient")) %>% 
  select(Parameter, estimate = mean, L95 = `2.5%`, U95 = `97.5%`, SE=sd) %>% 
  mutate(Region = "All Regions", Parameter = "BPP", Test = "PuMBA")

simbafit <- readRDS("../DerivedData/simdat_simbafit.rds")

simcompare_simba <- as_tibble(rstan::summary(simbafit)$summary, 
                           rownames = "Parameter") %>% 
  filter(Parameter=="b_logBPnd[10]") %>% 
  select(Parameter, estimate = mean, L95 = `2.5%`, U95 = `97.5%`, SE=sd) %>% 
  mutate(Region = "All Regions", Parameter = "BPP", Test = "SiMBA")
  
simcompare <- bind_rows(simcompare_t, 
                        simcompare_lme,
                        simcompare_pumba,
                        simcompare_simba) %>% 
  mutate(Comparison = paste0(Test, ": ", Region)) %>% 
  ungroup() %>% 
  mutate(Test = fct_inorder(Test),
         Comparison = fct_rev(fct_inorder(Comparison)))
```
Now let's plot the estimated group differences. Recall that the true value is 0.182, or 20%.

```{r}
ggplot(simcompare, aes(x=Comparison, y=estimate, colour=Test)) +
  geom_point() + 
  geom_errorbar(aes(ymin = L95, 
                    ymax = U95), width=0.2) +
  geom_hline(yintercept=0.182, colour="red", linetype="dashed") +
  coord_flip() +
  labs(
    title = "Model Group Difference Summaries : Log Scale",
    subtitle = "Difference Estimates and their 95% Confidence/Credible Intervals\nwith the true difference indicated by the dashed red line",
    y="Estimate (log scale)", x="Test"
  ) +
  scale_colour_brewer(type="qual", palette = 2)
```

Let's also compare the standard errors, describing the precision of our estimates:

```{r}
ggplot(simcompare, aes(x=Comparison, y=SE, colour=Test)) +
  geom_point(size=2.5, colour="black") + 
  geom_point(size=2) + 
  coord_flip() +
  labs(
    title = "Model Group Difference Standard Error",
    subtitle = "Standard Error of Difference Estimates: Lower is More Precise",
    y="Standard Error of the log-scale Estimates", x="Test"
  ) +
  scale_colour_brewer(type="qual", palette = 2)
```

Note that while SiMBA does outperform PuMBA, it is approximately 4000 times slower to estimate, restricted to the two-tissue compartment model, and requires defining more priors.


# Other Models

Having shown how PuMBA can be applied to data originating from the 2TC above, here I will show how PuMBA can be applied to data originating from other models: the one-tissue compartment model and the simplified reference tissue model.  In both cases, I will model the data using both PuMBA and LME for comparison.  Look for the row with `GroupPatient` to see the estimated difference, and compare the standard error between the two methods: for PuMBA, it is listed under `Est.Error` and for LME, it is `Std.Error`.

## SRTM: [11C]DASB

```{r}
srtm_modeldat <- readRDS("../RawData/DASB_SRTM_pars.rds")

head(srtm_modeldat)
```

This is a dataset of simulated parameters from a set of subjects belonging to a control and patient group, representing [11C]DASB data.  There is a difference of 0.1 in logBP~ND~ between the patients and the controls.

```{r}
ggplot(srtm_modeldat, aes(x=Group, y=logBPnd, colour=Group)) +
  geom_beeswarm(cex = 3) +
  facet_wrap(~Region)
```

Now let's model this with PuMBA

```{r}
bf_R1 <- bf(logR1 ~ 1 + Region + (1|k|ID), center = F)
bf_k2prime <- bf(logk2prime ~ 1 + (1|j|Region) + (1|k|ID), center = F)
bf_BPnd <- bf(logBPnd ~  1 + Region + Group + (1|k|ID))
```


```{r}

pumba_srtm_prior <- c(
  set_prior("normal( 0, 0.25)", resp = "logR1"),
  set_prior("normal(-3, 0.25)", resp = "logk2prime"),
  set_prior("normal(-1, 0.25)", resp = "logBPnd"),
  
  set_prior("normal(0, 0.3)", resp = "logR1", class = "sd", group="ID"),
  set_prior("normal(0, 0.3)", resp = "logk2prime", class = "sd", group="ID"),
  set_prior("normal(0, 0.3)", resp = "logBPnd", class = "sd", group="ID"),
  
  set_prior("normal(0, 0.1)", resp = "logk2prime", class = "sd", group="Region"),
  
  set_prior("normal(0, 0.3)", coef="Regionamy",     resp="logR1"),
  set_prior("normal(0, 0.3)", coef="Regioncin",     resp="logR1"),
  set_prior("normal(0, 0.3)", coef="Regiondpu",     resp="logR1"),
  set_prior("normal(0, 0.3)", coef="Regionhip",     resp="logR1"),
  set_prior("normal(0, 0.3)", coef="Regionins",     resp="logR1"),
  set_prior("normal(0, 0.3)", coef="Regionmidcs_c", resp="logR1"),
  set_prior("normal(0, 0.3)", coef="Regiontha",     resp="logR1"),
  set_prior("normal(0, 0.3)", coef="Regionvst",     resp="logR1"),

  set_prior("normal(1, 0.3)", coef="Regionamy",     resp="logBPnd"),
  set_prior("normal(0, 0.3)", coef="Regioncin",     resp="logBPnd"),
  set_prior("normal(1, 0.3)", coef="Regiondpu",     resp="logBPnd"),
  set_prior("normal(0, 0.3)", coef="Regionhip",     resp="logBPnd"),
  set_prior("normal(0.5, 0.3)", coef="Regionins",     resp="logBPnd"),
  set_prior("normal(1.5, 0.3)", coef="Regionmidcs_c",   resp="logBPnd"),
  set_prior("normal(1, 0.3)", coef="Regiontha",     resp="logBPnd"),
  set_prior("normal(1, 0.3)", coef="Regionvst",     resp="logBPnd"),
  
  set_prior("normal(0, 0.2)", coef="GroupPatient",     resp="logBPnd"),
  
  set_prior("normal(0, 0.1)", resp = "logR1", class = "sigma"),
  set_prior("normal(0, 0.1)", resp = "logk2prime", class = "sigma"),
  set_prior("normal(0, 0.1)", resp = "logBPnd", class = "sigma"),
  
  set_prior("lkj(1)", class="cor", group = "ID"),
  set_prior("lkj(2)", class="rescor"))


pumba_srtm <- brms::brm(formula = bf_R1 + bf_k2prime + bf_BPnd + set_rescor(TRUE), 
          data = srtm_modeldat, 
          family = gaussian(), 
          prior = pumba_srtm_prior, 
          cores=4, iter=4000,
          silent = 2, refresh=0, seed = 42)

pumba_srtm
```

And here is the LME model for comparison.

```{r}
summary(lmer(logBPnd ~  1 + Region + Group + (1|ID), data=srtm_modeldat))
```


## 1TC: [11C]GR103545

```{r}
onetc_modeldat <- readRDS("../RawData/GR_1TC_pars.rds") %>% 
  mutate(Region = relevel(as.factor(Region), ref = "dor"))

head(onetc_modeldat)
```

This is a dataset of simulated parameters from a set of subjects belonging to a control and patient group, representing [11C]GR103545 data.  There is a difference of 0.1 in logV~T~ between the patients and the controls.

```{r}
ggplot(onetc_modeldat, aes(x=Group, y=logVt, colour=Group)) +
  geom_beeswarm(cex = 3) +
  facet_wrap(~Region)
```

Now let's model this with PuMBA

```{r}
bf_K1 <- bf(logK1 ~ 1 + Region + (1|k|ID), center = F)
bf_Vt <- bf(logVt ~ 1 + Region + Group + (1|k|ID), center = F)
```


```{r}
pumba_onetc_prior <- c(
  set_prior("normal(-2, 0.25)", resp = "logK1", coef="Intercept", class="b"),
  set_prior("normal(3, 0.25)", resp = "logVt", coef="Intercept", class="b"),
  
  set_prior("normal(0, 0.3)", resp = "logK1", class = "sd", group="ID"),
  set_prior("normal(0, 0.3)", resp = "logVt", class = "sd", group="ID"),
  
  set_prior("normal(0, 0.1)", resp = "logK1", class = "sigma"),
  set_prior("normal(0, 0.1)", resp = "logVt", class = "sigma"),
  
  set_prior("normal(0, 0.3)", coef="Regionamy",         resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regiondpu",         resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regionhip",         resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regionins",         resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regionmed",         resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regionpip",         resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regionraphe_t_all", resp="logK1"),
  set_prior("normal(0, 0.3)", coef="Regionvst",         resp="logK1"),

  set_prior("normal(0, 0.3)", coef="Regionamy",         resp="logVt"),
  set_prior("normal(0, 0.3)", coef="Regiondpu",         resp="logVt"),
  set_prior("normal(0, 0.3)", coef="Regionhip",         resp="logVt"),
  set_prior("normal(0, 0.3)", coef="Regionins",         resp="logVt"),
  set_prior("normal(0, 0.3)", coef="Regionmed",         resp="logVt"),
  set_prior("normal(0, 0.3)", coef="Regionpip",         resp="logVt"),
  set_prior("normal(0, 0.3)", coef="Regionraphe_t_all", resp="logVt"),
  set_prior("normal(0, 0.3)", coef="Regionvst",         resp="logVt"),

  set_prior("normal(0, 0.2)", coef="GroupPatient", resp="logVt"),
  
  set_prior("lkj(1)", class="cor", group = "ID"),
  set_prior("lkj(2)", class="rescor"))


pumba_1tc <- brms::brm(formula = bf_K1 + bf_Vt + set_rescor(TRUE), 
          data = onetc_modeldat, 
          family = gaussian(), 
          prior = pumba_onetc_prior, 
          cores=4, iter=4000,
          silent = 2, refresh=0, seed = 42)

pumba_1tc
```

And here is the LME model for comparison.

```{r}
summary(lmer(logVt ~  1 + Region + Group + (1|ID), data=onetc_modeldat))
```

